---
title: "GFSI-Time-Series"
author: "SETH-SILVER"
date: "10/9/2020"
output: html_document
---

# Imorting the relevant libraries
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(gganimate)
library(gifski)
```

## Importing datasets
```{r}
d2016 <- read_csv('GFSI-2016.csv')
d2017 <- read_csv('GFSI-2017.csv')
d2018 <- read_csv('GFSI-2018.csv')
d2019 <- read_csv('GFSI-2019.csv')
d2015 <- read_csv('GFSI-2015.csv')
```

## Removing Redudant (X1) columns
```{r}
d2015 <- d2015 %>% select(-X1)
d2016 <- d2016 %>% select(-X1)
d2017 <- d2017 %>% select(-X1)
d2018 <- d2018 %>% select(-X1)
d2019 <- d2019 %>% select(-X1)
```


## Selecting some low income countries from the datasets
```{r}
ld2015 = d2015 %>% filter(Series %in% c('Uganda','Ethiopia','Burundi','Benin','Mozambique',
                                      'Mali','Malawi','Rwanda'))
ld2016 = d2016 %>% filter(Series %in% c('Uganda','Ethiopia','Burundi','Benin','Mozambique',
                                        'Mali','Malawi','Rwanda'))
ld2017 = d2017 %>% filter(Series %in% c('Uganda','Ethiopia','Burundi','Benin','Mozambique',
                                        'Mali','Malawi','Rwanda'))
ld2018 = d2018 %>% filter(Series %in% c('Uganda','Ethiopia','Burundi','Benin','Mozambique',
                                        'Mali','Malawi','Rwanda'))
ld2019 = d2019 %>% filter(Series %in% c('Uganda','Ethiopia','Burundi','Benin','Mozambique',
                                        'Mali','Malawi','Rwanda'))
```


## Appending a 'Year' Column to each dataset respectively
```{r}
ld2015$Year <- c(2015, 2015, 2015, 2015, 2015, 2015, 2015, 2015)
ld2016$Year <- c(2016, 2016, 2016, 2016, 2016, 2016, 2016, 2016)
ld2017$Year <- c(2017, 2017, 2017, 2017, 2017, 2017, 2017, 2017)
ld2018$Year <- c(2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018)
ld2019$Year <- c(2019, 2019, 2019, 2019, 2019, 2019, 2019, 2019)
```


## Selecting desired variables and binding the data
```{r}
ld2015 <- ld2015 %>% select(`Series`, `Year`, `1.1) Change in average food costs`, `1.3) Gross domestic product per capita (US$ PPP)`,
                            `3.1) Dietary diversity`, `2.8) Food loss`, `2.7) Urban absorption capacity`, `1.4) Agricultural import tariffs`,
                            `2.1.1) Average food supply`, `2.4) Volatility of agricultural production`, `2.5) Political stability risk`,
                            `2.6) Corruption`, `4.1.6) Commitment to managing exposure`,`4.5.1) Food import dependency`, `4.5.2) Dependence on natural capital`,
                            `4.5.3) Disaster risk management`)
ld2016 <- ld2016 %>% select(`Series`,`Year`, `1.1) Change in average food costs`, `1.3) Gross domestic product per capita (US$ PPP)`,
                            `3.1) Dietary diversity`, `2.8) Food loss`, `2.7) Urban absorption capacity`, `1.4) Agricultural import tariffs`,
                            `2.1.1) Average food supply`, `2.4) Volatility of agricultural production`, `2.5) Political stability risk`,
                            `2.6) Corruption`, `4.1.6) Commitment to managing exposure`,`4.5.1) Food import dependency`, `4.5.2) Dependence on natural capital`,
                            `4.5.3) Disaster risk management`)
ld2017 <- ld2017 %>% select(`Series`,`Year`, `1.1) Change in average food costs`, `1.3) Gross domestic product per capita (US$ PPP)`,
                            `3.1) Dietary diversity`, `2.8) Food loss`, `2.7) Urban absorption capacity`, `1.4) Agricultural import tariffs`,
                            `2.1.1) Average food supply`, `2.4) Volatility of agricultural production`, `2.5) Political stability risk`,
                            `2.6) Corruption`, `4.1.6) Commitment to managing exposure`,`4.5.1) Food import dependency`, `4.5.2) Dependence on natural capital`,
                            `4.5.3) Disaster risk management`)
ld2018 <- ld2018 %>% select(`Series`,`Year`, `1.1) Change in average food costs`, `1.3) Gross domestic product per capita (US$ PPP)`,
                            `3.1) Dietary diversity`, `2.8) Food loss`, `2.7) Urban absorption capacity`, `1.4) Agricultural import tariffs`,
                            `2.1.1) Average food supply`, `2.4) Volatility of agricultural production`, `2.5) Political stability risk`,
                            `2.6) Corruption`, `4.1.6) Commitment to managing exposure`,`4.5.1) Food import dependency`, `4.5.2) Dependence on natural capital`,
                            `4.5.3) Disaster risk management`)
ld2019 <- ld2019 %>% select(`Series`,`Year`, `1.1) Change in average food costs`, `1.3) Gross domestic product per capita (US$ PPP)`,
                            `3.1) Dietary diversity`, `2.8) Food loss`, `2.7) Urban absorption capacity`, `1.4) Agricultural import tariffs`,
                            `2.1.1) Average food supply`, `2.4) Volatility of agricultural production`, `2.5) Political stability risk`,
                            `2.6) Corruption`, `4.1.6) Commitment to managing exposure`,`4.5.1) Food import dependency`, `4.5.2) Dependence on natural capital`,
                            `4.5.3) Disaster risk management`)

data <- rbind(ld2015, ld2016, ld2017, ld2018, ld2019)
```


# PLOTTING THE DATA
# 1. Average price changes
```{r}
ggplot(data = data, mapping = aes(x = `Year`, y = `1.1) Change in average food costs`,
                                  fill = `Series`))+
  geom_col()+
  ggtitle('Average Price Changes; 2015-2019')
ggplot(data = data, mapping = aes(x = `Year`, y = `1.1) Change in average food costs`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Average Price Changes; 2015-2019')

#-----------------ANIMATED LINE PLOT: Uncomment to run if preferred------------------------------
#w <- ggplot(data = data, mapping = aes(x = `Year`, y = `1.1) Change in average food costs`,
#                                  color = `Series`))+
#  geom_line()+
#  ggtitle('Average Price Changes; 2015-2019')
#w + transition_reveal(`Year`)
```


# 2------Gross Domestic product per capita---------
```{r}
#expresses the country's income, as distributed to the entire population [GDP/Population]
ggplot(data = data, mapping = aes(x = `Year`, y = `1.3) Gross domestic product per capita (US$ PPP)`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Gross Domestic Product per capita')
```


#-------------------------Dietary Diversity-------------------------
```{r}
# changes in percentage dietary diversity
ggplot(data = data, mapping = aes(x = `Year`, y = `3.1) Dietary diversity`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Percentage Dietary Diversity')

ggplot(data = data, mapping = aes(x = `Year`, y = `3.1) Dietary diversity`,
                                  fill = `Series`))+
  geom_col()
```

#Food Loss
```{r}
#------------------Total waste/Total Domestic Supply---------------------------------------
ggplot(data = data, mapping = aes(x = `Year`, y = `2.8) Food loss`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Food Loss(Tonnes) [Total waste/Domestic Supply]')
```

#--------------------Urban Absorption Capacity------------------------------------------
```{r}
# %GDP change - Urban growth rate
ggplot(data = data, mapping = aes(x = `Year`, y = `2.7) Urban absorption capacity`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Urban Absorption Capacity [%GDP Change - Urban Growth Rate]')
```


#---------------------Agric Import Tariffs-----------------------------------------------
```{r}
# Low import tariffs may imply low internal agricultural production, and therefore an implied need
#to get products without, by importation, and the reverse is true
ggplot(data = data, mapping = aes(x = `Year`, y = `1.4) Agricultural import tariffs`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Agric Import Tariffs')
#ggplot(data = data, mapping = aes(x = `Year`, y = `1.4) Agricultural import tariffs`, fill = `Series`))+
#  geom_col()
```

#----------------Average food supply---------------------------------------------
```{r}
# Plot marred by missing values thus hardly important
ggplot(data = data, mapping = aes(x = `Year`, y = `2.1.1) Average food supply`))+
  geom_line()
  ggtitle('Average food supply')
ggplot(data = data, mapping = aes(x = `Year`, y = `2.1.1) Average food supply`, fill = `Series`))+
  geom_col()
```

#-----------------------Volatility in agricultural production----------------------------------------
```{r}
# These can imply increase or decrease in production
ggplot(data = data, mapping = aes(x = `Year`, y = `2.4) Volatility of agricultural production`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Volatility In Agric Production [S.D]')
ggplot(data = data, mapping = aes(x = `Year`, y = `2.4) Volatility of agricultural production`,
                                  fill = `Series`))+
  geom_col()+
  ggtitle('Volatility In Agric Production [S.D]')
```


#-----------------------Political Stability Risk--------------------------------------------------------
```{r}
ggplot(data = data, mapping = aes(x = `Year`, y = `2.5) Political stability risk`,
                                  color = `Series`))+
  geom_smooth()+
  ggtitle('Political Stability Risk (%)')
ggplot(data = data, mapping = aes(x = `Year`, y = `2.5) Political stability risk`,
                                  fill = `Series`))+
  geom_col()+
  ggtitle('Political Stability Risk (%)')
```


#------------------Corruption [Ranked over four]-----------------------------------------------------
```{r}
ggplot(data = data, mapping = aes(x = `Year`, y = `2.6) Corruption`))+
  facet_wrap(~`Series`)+
  geom_col(na.rm = TRUE)+
  ggtitle('Corruption Rank [4 = Highest]')

ggplot(data = data, mapping = aes(x = `Series`, y = `2.6) Corruption`, fill = `Series`))+
  facet_wrap(~`Year`)+
  coord_flip()+
  geom_col(na.rm = TRUE)+
  ggtitle('Corruption Rank [4 = Highest]')
```

#---------------------Managing Exposure--------------
```{r}
#Has lots of missing values: changes and corrections will be made in the near future
ggplot(data = data, mapping = aes(x = `Year`, y = `4.1.6) Commitment to managing exposure`,
                                  color = `Series`))+
  geom_smooth(na.rm = TRUE)
```

